### 2024-07-08 paper/paper.qmd
updating heatmap

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Sehar Bajwa
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
prefer-html: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#install.packages("usethis")
#install.packages("gitcreds")
#library(usethis)
#library(gitcreds)
#create_github_token()
# Example dataframe with values and ranges
data_ranges <- data.frame(
  category = c("A", "B", "C", "D"),
  value = c(100, 150, 200, 250),
  range = c(10, 15, 20, 25)  # Example range values
)

# Calculate upper and lower bounds for plotting
data_ranges$lower <- data_ranges$value - data_ranges$range
data_ranges$upper <- data_ranges$value + data_ranges$range

# Print the dataframe
print("Data with Ranges:")
print(data_ranges)
# Load necessary library
library(ggplot2)

# Create the plot
ggplot(data_ranges, aes(x = category, y = value)) +
  geom_point() +  # Plot points for values
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +  # Error bars for ranges
  labs(title = "Values with Ranges", x = "Category", y = "Value") +
  theme_minimal()

library(tidyverse)
library(palmerpenguins)
```

# Introduction

In the biggest democratic election the world has ever seen, India held its general election in 2024 over 7 successive phases, beginning on the 19th of April and ending on the 1st of June. The election determined the composition of the Lok Sabha, the lower house of India's Parliament, which comprises 543 seats. A party or coalition needs a minimum of 272 seats to form a government.

During the election period, numerous exit and opinion polls were conducted to gauge public sentiment and predict the potential outcome. Opinion polls were conducted in the months leading up to the election, providing insights into voter preferences and trends. Exit polls were conducted immediately after voters cast their ballots, offering a snapshot of the electorate's choices.

The results of these polls were keenly watched by political analysts, parties, and the public, as they provided an early indication of which party or coalition might secure a majority in the Lok Sabha. Despite their limitations and varying degrees of accuracy, these polls played a crucial role in shaping the narrative and expectations surrounding the 2024 Indian general election.

# Background  {#sec-background}

Before delving into the following sections of the paper, it is important to provide some political context. 
The National Democratic Alliance (NDA) is a centre-right to right wing political alliance, led by the Bharatiya Janata Party (BJP), and has been a significant political force in India since its formation in the late 1990s, particularly under Prime Minister Narendra Modi, who assumed office in 2014 and was re-elected in 2019.
In 2023, the Indian National Developmental Inclusive Alliance (INDIA) was established, comprising major opposition parties from across the political spectrum, including the Indian National Congress (INC), the Dravida Munnetra Kazhagam (DMK), and the Trinamool Congress (TMC). This centre-left to left wing coalition, headed by Rahul Gandhi of the INC, was formed to present a united front against the BJP's dominance, driven by concerns over Modi's prolonged tenure. The INDIA Alliance's collaboration aimed to make a dent in the NDA's mammoth vote bank by consolidating votes and pooling in resources. And they did achieve the goal of inhibiting the NDA's expected landslide victory (353 seats in 2019). The slogans of '400 paar' (crossing 400 seats) by the NDA this time around were paid a lot of heed, and INDIA's alliance was underestimated by the opinion polls, and perhaps falsely represented by the exit polls.

The newshouses received a lot of flak for the inflation of the exit poll numbers in particular, nfluenced by the BJP's past electoral success and media narrative. In addition to misleading the public, they led to the worst stock market crash in the country in the past 4 years.The term "godi media" colloquially describes Indian media outlets perceived as biased towards the BJP, often criticized for their lack of independent journalism.

Consequently, the analysis in this paper will focus on the National Democratic Alliance (NDA) because of the central role it played in the 2024 Indian general election. With Prime Minister Narendra Modi seeking a third consecutive term, the election was largely seen as a referendum on his leadership and the performance of his Bharatiya Janata Party (BJP). The political discourse and public sentiment were heavily polarized around the NDA, with debates and predictions revolving around whether Modi and the BJP could maintain their stronghold on Indian politics. Consequently, all bets and polls were discussions around the electoral landscape being pro or anti-NDA, instead of being pro INDIA or pro Independent candidates.

# Data {#sec-data}

The data used in this paper is obtained by scraping the Wikipedia page for 'general elections in India 2024', and cross-referenced with each article it mentioned. There seems to be no singular database to collect all pre-election and exit poll data, and this is part of the gaping problem examined in this paper. The data points are determined from several National and local newspapers that have published the news online.

While the results, possible margins, and sample sizes for opinion polls are available online, detailed reports from the research or polling agencies hired by news outlets are not publicly accessible on their websites.

The three data sets I will be using are Opinion Poll data, Exit Poll data and Mood of the Nation (CVoter).

## Opinion Poll data

In the run-up to the 2024 General Election, various newspapers and agencies conduct pre-election polls to gauge the mindset, mood, and political leaning of the average voter. These polls provide valuable insights into public opinion, allowing political analysts, parties, and the general public to understand potential election outcomes better. They help identify trends, shifts in voter preferences, and the impact of current events and policies on voter behavior.

Containing data from January 2023 up to April 2024, this data consists of the following 7 variables: name of the polling agency, sample size, seats won by NDA, INDIA, and Others. Additionally, the 'Exact' date published is a self-created variable after scouring the references, so a more accurate dataset can be made, and the overlap in a line graph can be avoided. A snapshot of the cleaned data is provided below, which includes the 7 variables and 17 data points in total.

```{r}
#| include: false
#| warning: false
#| message: false

#### Initialize quarto workspace by reading csv's ####
library(tidyverse)
library(here)
library(readr)
library(ggplot2)

cleaned_opinion_polldata <- read_csv(here::here("/cloud/project/data/analysis_data/cleaned_opinion_poll.csv"))
cleaned_exit_polldata <- read_csv(here::here("/cloud/project/data/analysis_data/cleaned_exit_poll.csv"))
MOTN_polldata <- read_csv(here::here("/cloud/project/data/raw_data/MOTN.csv"))

```

```{r}
#| echo: false
#| warning: false
#| label: tbl-dataa
#| tbl-cap: "A summary atable of cleaned data"
#install.packages("knitr")
#install.packages("kableExtra")
library(kableExtra)
library(knitr)
kable(cleaned_opinion_polldata, padding = 2)%>%
  kable_styling(font_size = 8) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4, width = "2cm") %>%
  column_spec(5, width = "1.5cm") %>%
  column_spec(6, width = "1.5cm") %>%
  column_spec(7, width = "1.5cm")

```

## Mood of the Nation data

Conducted by National Newshouse India Today, the CVoter opinion poll utilizes Computer Assisted Telephone Interviewing among voters aged 18 and above across all 543 Lok Sabha seats. Targeting a sample of 30,000 per quarter with an average response rate of 55%, the tracker is conducted weekly in 11 national languages. According to the official site, CVoter employs a Standard Resilient Distributed dataset to randomly select numbers across telecom lines, ensuring a representative analysis through statistical weighting of the data to reflect the local population as per the latest Census figures. The data is weighted to align with known census profiles, including gender, age, education, income, religion, caste, urban/rural distribution, and past vote recall for the Lok Sabha and Vidhan Sabha elections. For analytics, CVoter uses a proprietary algorithm to calculate provincial and regional vote shares based on the split-voter phenomenon.

The CVoter data offers the most consistent output for opinion polls in India and has been selected due to its standardization and regularity. This consistency allows for the assumption that all other variables are controlled, enabling a focused analysis of trends. The dataset has been independently constructed by aggregating infographics and articles published on the India Today website. The included dataset only has 5 variables, the month, and NDA, INDIA and Others Seats. A snapshot is provided below.

```{r}
#| echo: false
#| warning: false
#| label: tbl-dDAata
#| tbl-cap: "A summary table of cleaned data"
#install.packages("knitr")
#install.packages("kableExtra")
library(kableExtra)
library(knitr)
kable(MOTN_polldata, padding = 2)%>%
  kable_styling(font_size = 8) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4, width = "2cm")

```

## Exit poll data

The inception of exit polls in India traces back to 1957, during the second Lok Sabha elections, with the pioneering initiative led by the Indian Institute of Public Opinion.

Exit polls, as the name suggests, are conducted after the culmination of the last phase of voting, in accordance with the guidelines set by the Election Commission of India. Distinguished from opinion polls, which precede elections, exit polls involve surveys administered to voters as they exit polling stations post casting their ballots. These surveys encompass a spectrum of inquiries, delving into voter motivations and party preferences.. As the 2024 Lok Sabha elections concluded, exit poll results for 28 states and eight union territories were slated for release from 6:30 pm onwards on June 1 2024, in compliance with Election Commission protocols.

In contrast to opinion polls, exit polls are often conducted and released hastily on the same day, rendering the timing irrelevant. Moreover, these exit polls have not disclosed the sample size of the respondents surveyed. As a result, they are presented here merely as indicators, highlighting their potential for inaccuracy and inflation. There are 4 variables, the name of the agency, and the number of seats predicted each for NDA, INDIA and others. The 15 datapoints have been scraped from Wikipedia using rvest.

```{r}
#| echo: false
#| warning: false
#| label: tbl-data
#| tbl-cap: "A summary table of dsdasa data"
#install.packages("knitr")
#install.packages("kableExtra")
library(kableExtra)
library(knitr)
kable(cleaned_exit_polldata, padding = 2)%>%
  kable_styling(font_size = 8) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4, width = "2cm")

```

# Model

The following section features line graphs and simple linear regression models to provide a closer look at each of the datasets in turn.

## Opinion poll data

By isolating all three party coalitions, it is easier to notice trends in the seat projections in the months leading up to the election. Particularly for the NDA, the predictions became skewed and deviated further from a more realistic end result. The target starts as somewhat realistic in April 2023, given that the NDA won 353 seats in the 2019 General Elections, and the general opinion of the public seems to be that they will surpass this previous record with the next election. However, as the election date approached, the projections increased dramatically, peaking at over 400 seats. This suggests an overestimation of NDA's popularity or a possible bias in the polling data.
Conversely, INDIA being a relatively young coalition, sees dynamic changes in predictions and a general underconfidence as the underdog in the election. Initially, the projections are relatively stable but show a significant drop around October 2023, which indicates possible public skepticism or fluctuation in support. The coalition's seat projections are consistently below the actual outcome, highlighting the tendency of the polls to underestimate their potential.
The predictions for the Others seats (which notably included the Bahujan Samaj Party and the AIADMK) were the most accurate, mostly because these are highly regional parties, making it easier to predict the seats they would bag with confidence. The projections for Others show a steady decline, aligning closely with the actual results. This accuracy can be attributed to the strong regional bases and consistent voter support for these parties, which makes their outcomes more predictable compared to the broader national coalitions.

```{r}
#| echo: false
#| warning: false
#| label: fig-opinionpoll
#| fig-cap: "Opinion poll Seat Projections"
#| fig-subcap: ["NDA Projections", "INDIA Projections", "Others Projections"]
#| layout-ncol: 2

# Load necessary libraries
library(tidyverse)
library(arrow)
library(ggplot2)
library(gridExtra)

# Step 2: Save the data as a Parquet file
opinionpoll_parquet <- "/cloud/project/data/analysis_data/cleaned_opinion_poll.parquet"
write_parquet(cleaned_opinion_polldata, opinionpoll_parquet)

# Step 3: Sort the data by 'Exact date' in descending order
opinion_poll_sorted <- cleaned_opinion_polldata %>%
  arrange(desc(Exact))

# Step 4: Create a line graph for NDA
plot1 <- ggplot(opinion_poll_sorted, aes(x = as.Date(Exact), y = NDA)) +
  geom_line(color = "blue") +
  geom_hline(yintercept = 293, linetype = "dashed", color = "orange") +
  annotate("text", x = as.Date("2024-01-01"), y = 300, label = "Actual NDA seats", color = "orange") +
  labs(x = "Date of publication", y = "NDA Seats", title = "Opinion poll-NDA Seat Projections") +
  theme_minimal()
print(plot1)

# Step 4: Create a line graph for INDIA
plot2 <- ggplot(opinion_poll_sorted, aes(x = as.Date(Exact), y = INDIA)) +
  geom_line(color = "blue") +
  geom_hline(yintercept = 234, linetype = "dashed", color = "lightblue") +
  annotate("text", x = as.Date("2024-01-01"), y = 240, label = "Actual INDIA seats", color = "lightblue") +
  labs(x = "Date of publication", y = "INDIA Seats", title = "Opinion poll-INDIA Seat Projections") +
  theme_minimal()
print(plot2)

# Step 4: Create a line graph for Others
plot3 <- ggplot(opinion_poll_sorted, aes(x = as.Date(Exact), y = Others)) +
  geom_line(color = "blue") +
  geom_hline(yintercept = 16, linetype = "dashed", color = "green") +
  annotate("text", x = as.Date("2024-01-01"), y = 25, label = "Actual Others seats", color = "green") +
  labs(x = "Date of publication", y = "Others Seats", title = "Opinion poll-Others Seat Projections") +
  theme_minimal()

print(plot3)
```

## Mood of the Nation

By controlling for variables such as sample size and methods of questioning (presented in section 1.2), the popularity and public opinion of the governing party can be examined in isolation. While the Mood of the Nation data does not follow a noticeable trend other than sloping upwards, the lowest points of the graph coincide with significant challenges during the ruling party's tenure. This is particularly evident from August 2021 to January 2022, where seat projections hit record lows around 296. Notably, this period aligns with the second wave of the Covid-19 pandemic, which severely overwhelmed India's healthcare system, leading to deficits in hospitals, medicines, and oxygen cylinders.

The pandemic exacerbated existing economic issues, such as high unemployment rates and a marked slowdown in economic activity. Rising inflation rates further complicated the situation, and in the first quarter of 2021, India experienced a 'mild' recession, with GDP dropping by a record 24%. Additionally, widespread dissatisfaction with the Farm Bills led to high-profile protests across the country, impacting India’s agriculture-dominant economy.

```{r}
#| echo: false
#| warning: false
#| label: fig-opienionpoll
#| fig-cap: "Opinion poll Seat Projections"

# Read the CSV file
motngraph_data <- read_csv("/cloud/project/data/raw_data/MOTN.csv")

# Convert the Month column from character to date format
motngraph_data <- motngraph_data %>%
  mutate(Month = as.Date(paste("01", Month, sep = "-"), format = "%d-%b-%y"))

# Save the dataframe as a Parquet file
write_parquet(motngraph_data, "/cloud/project/data/analysis_data/MOTN.parquet")

# Create a line plot for NDA seats over time
plot_motn <- ggplot(motngraph_data, aes(x = Month, y = NDA)) +
  geom_line(color = "blue") +
  geom_point() +
  geom_text(aes(label = NDA), vjust = -0.5, color = "black", size = 3) +
  labs(x = "Month", y = "NDA Seats", title = "NDA Seats Over Time") +
  theme_minimal() +
  geom_hline(yintercept = 293, linetype = "dashed", color = "red") +
  annotate("text", x = as.Date("2023-01-01"), y = 295, label = "Actual NDA seats", color = "red") +
  scale_x_date(date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Print the plot
print(plot_motn)

```

It is generally understood that when the sample size in a study increases, it represents a larger subset of the population and should therefore be more accurate about the average since it reduces sample-to-sample variability. Even if polls are conducted by different agencies, it is reasonable to assume that larger sample sizes should yield more accurate seat projections.

To test this assumption, a linear regression model was created using rstanarm to see if there is a correlation between sample size and the accuracy of the opinion poll predictions. One particular data point stood out as an outlier due to its exceptionally large sample size: the TV9-Bharatvarsh poll conducted in April 2024, with a sample size of 2,500,000 voters. A graph excluding this point was created to better visualize the data points clustering between 0 and 500,000.

The results showed no significant correlation between sample size and the accuracy of seat projections. The p-value was 0.6, which is far greater than the standard threshold of 0.05 for statistical significance. Interestingly, the trendline suggested that smaller sample sizes were more accurate in their predictions, contradicting the initial assumption. In other words, as more participants were included to represent a subset of the population more accurately, the results skewed and varied further away from the true value. A few reasons for the same are discussed in section .

```{r}
#| echo: false
#| warning: false
#| label: fig-oprrinionpoll
#| fig-cap: "Opinion poll Seat Projections"
#| #| fig-subcap: ["NDA Projections", "INDIA Projections"]
#| layout-ncol: 2

opinion_lm <- read_csv("/cloud/project/data/analysis_data/cleaned_opinion_poll.csv")

# Fit a linear regression model
lm_model <- lm(NDA ~ `Sample size`, data = opinion_lm)

# Print summary of the linear model
#summary(lm_model)

# Extract coefficients and their statistics
summary_table <- summary(lm_model)$coefficients %>%
  as.data.frame() %>%
  rownames_to_column(var = "Coefficient") %>%
  rename(estimate = Estimate, std_error = `Std. Error`, t_value = `t value`, p_value = `Pr(>|t|)`)

# Print the summary table
#print(summary_table)

# Check if the relationship is statistically significant (p-value < 0.05)
#p_value <- summary(lm_model)$coefficients["`Sample size`", "Pr(>|t|)"]
#if (!is.na(p_value) && p_value < 0.05) {
#  cat("The relationship between Sample size and NDA seats is statistically significant (p < 0.05).\n")
#} else {
#  cat("The relationship between Sample size and NDA seats is not statistically significant (p >= 0.05).\n")
#}

# Plot scatter plot with regression line
plot_opinion_poll <- ggplot(opinion_lm, aes(x = `Sample size`, y = NDA)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_hline(yintercept = 293, linetype = "dashed", color = "red") +
  annotate("text", x = 1000000, y = 300, label = "Actual NDA seats", color = "red") +
  labs(x = "Sample Size", y = "NDA Seats", title = "Sample Size vs. NDA Seats with Linear Regression") +
  theme_minimal()

# Print the plot
print(plot_opinion_poll)

filteredopinion_lm <- read_csv("/cloud/project/data/analysis_data/cleaned_opinion_poll.csv")

# Filter out rows where Sample size is 2500000
opinion_poll_filtered <- filteredopinion_lm %>%
  filter(`Sample size` != 2500000)

# Fit a linear regression model
lm_model <- lm(NDA ~ `Sample size`, data = opinion_poll_filtered)

# Print summary of the linear model
#summary(lm_model)

# Extract coefficients and their statistics
summary_table <- summary(lm_model)$coefficients %>%
  as.data.frame() %>%
  rownames_to_column(var = "Coefficient") %>%
  rename(estimate = Estimate, std_error = `Std. Error`, t_value = `t value`, p_value = `Pr(>|t|)`)

# Print the summary table
#print(summary_table)

# Check if the relationship is statistically significant (p-value < 0.05)
#p_value <- summary(lm_model)$coefficients["`Sample size`", "Pr(>|t|)"]
#if (!is.na(p_value) && p_value < 0.05) {
#  cat("The relationship between Sample size and NDA seats is statistically significant (p < 0.05).\n")
#} else {
 # cat("The relationship between Sample size and NDA seats is not statistically significant (p >= 0.05).\n")
#}

# Plot scatter plot with regression line (excluding Sample size == 2500000)
plot_opinion_poll_filtered <- ggplot(opinion_poll_filtered, aes(x = `Sample size`, y = NDA)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_hline(yintercept = 293, linetype = "dashed", color = "red") +
  annotate("text", x = 100000, y = 300, label = "Actual NDA seats", color = "red") +
  labs(x = "Sample Size", y = "NDA Seats", title = "Sample Size vs. NDA Seats (Excluding 2500000)") +
  theme_minimal()

# Print the plot
print(plot_opinion_poll_filtered)
```

## The exit polls and godi media

As mentioned in previous sections, 'godi media' and inflated opinion poll numbers lead to uncharacteristically high predictions for a right-wing landslide victory in India. After scouring multiple media articles, social media opinions, and analyzing general prevalent public opinions through conversations (compiled in the resources folder in __), a pro-right rating can be collated for the media houses. Here is an excerpt from one such source:

Prominent among these pro-BJP/RSS news outlets are English television channels Republic TV, India Today, Wion, and CNN-News18, and Hindi TV channels Zee News, India TV, AajTak, ABP News, Republic Bharat, Sudershan News, News Nation, and News24 (India).

Conversely, DB Live is widely understood to be what many refer to as a 'mouthpiece of the Indian National Congress'. It is one of the few channels in a rapidly declining list to show outward criticism of the ruling government's policies. Out of 180 countries, India ranked 161 in the Reporters Without Borders' World Press Freedom index in 2023, down 21 positions since Modi became prime minister in 2014. According to media watchdog NewsLaundry, an analysis of 429 primetime segments by 6 anchors in the run-up to the polls revealed that 52% of on-air time was spent criticizing the opposition, while only a small percentage of that time was spent on topics like jobs, education, and social issues.

Heatmap Analysis
To visually represent the relationship between the pro-NDA bias and seat projections, a heatmap of the Pro-NDA score alongside the seat projections of various news agencies is presented below. This graph should be approached with caution due to minimal secondary research available to support it, but it provides a valuable lens to consider when examining media bias and its potential impact on opinion poll results.

```{r}
#| echo: false
#| warning: false
#| label: fig-oprrinionffpoll
#| fig-cap: "Opinion poll Seat Prosdfjections"

media_data <- data.frame(
  Media_Outlets = c("ABP News-CVoter", "Agni News Services", "Dainik Bhaskar", "DB Live", 
                    "India Today-Axis My India", "India News-Dynamics", "India TV-CNX", 
                    "NDTV-Jan Ki Baat", "News18-CNBC", "News 24-Today's Chanakya", 
                    "News Nation", "Republic TV-Matrize", "Republic TV-PMarq", 
                    "Times Now-ETG", "TV9 Bharatvarsh"),
  NDA_Predictions = c(368, 242, 316, 221, 381, 371, 386, 377, 362, 400, 360, 360, 359, 358, 346),
  Pro_BJP_Score = c(7, 2, 4, 1, 9, 6, 8, 7, 5, 10, 3, 9, 9, 8, 7)  
)

# Sort data by Pro_BJP_Score
media_data <- media_data[order(media_data$Pro_BJP_Score, decreasing = TRUE), ]

# Create the heat map
ggplot(media_data, aes(x = reorder(Media_Outlets, Pro_BJP_Score), y = "NDA Prediction")) +
  geom_tile(aes(fill = NDA_Predictions), color = "white") +
  scale_fill_gradient(low = "blue", high = "orange", name = "NDA Predictions") +
  labs(title = "Heat Map of NDA Seat Predictions by Media Outlet",
       x = "Media Outlets",
       y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

\newpage

# References